---
title: "STAT 491 - Pokémon Project Code"
author: "Harley Clifton & Jordyn Lucier"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
library(rjags)
knitr::opts_chunk$set(echo = TRUE)
```

Using the dataset for your project, complete the following tasks. Note this mirrors the second question on the takehome exam.

#### Research Question

**How does Pokemon type impact overall stats? In other words, which Pokemon types tend to have the highest stats on average?**


$~$

#### Exploratory Data Analysis


```{r, out.width="450px", fig.align='center'}
# Data Wrangling
pokemon.dat <- read_csv("data/Pokemon_2.0.csv")

names <- c("Type", "Type2", "Generation", "Evolution")
pokemon.dat[,names] <- lapply(pokemon.dat[ , names], factor)

summary(pokemon.dat)

levels(pokemon.dat$Type) # there are 18 pokemon types


# Data Visualization
ggplot(data = pokemon.dat, aes(x = Type, y = Total, fill = Type)) +
  geom_violin() +
  geom_jitter(width = 0.1, alpha = 0.3) +
  labs(title = "Pokemon Total Stats by Type",
       x = "Pokemon Type",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.position = "none")

# c("maroon1", "darkslateblue", "mediumvioletred", "paleturquoise", "thistle1",
# "sienna1", "red", "lightskyblue", "seashell1", "lightgreen", "peachpuff3",
# "powderblue", "orange1", "darkmagenta", "plum2",  "lightsalmon4", "lightgrey",
# "mediumturquoise")
```

```{r}
nrow(pokemon.dat)
count(pokemon.dat$Type)

count(pokemon.dat$Type2)

count(pokemon.dat$Generation)

count(pokemon.dat$Evolution)
aggregate(pokemon.dat$Total~pokemon.dat$Evolution, FUN=mean)

count(pokemon.dat$Legendary)
```


```{r}
gen.colors <- c("rosybrown1", "pink1", "palevioletred1", 
                "maroon1",  "maroon2", "mediumvioletred")

ggplot(data = pokemon.dat, aes(x = Generation, y = Total)) +
  geom_violin(aes(fill = Generation)) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  scale_fill_manual(values = gen.colors) +
  labs(title = "Pokemon Total Stats by Generation",
       x = "Generation",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5),
        legend.position = "none") 
```



```{r}
evol.colors <- c("burlywood1", "goldenrod1", "orange1", "darkorange2")

ggplot(data = pokemon.dat, aes(x = Evolution, y = Total)) +
  geom_violin(aes(fill = Evolution)) +
  geom_jitter(width = 0.1, alpha = 0.1) +
  scale_fill_manual(values = evol.colors) +
  labs(title = "Pokemon Total Stats by Evolution",
       x = "Evolution",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5),
        legend.position = "none")


ggplot(data = pokemon.dat, aes(x = Evolution, y = Total)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm, color = "goldenrod1") +
  geom_smooth(method = loess, color = "darkorange2") +
  labs(title = "Pokemon Total Stats by Evolution",
       x = "Evolution",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5),
        legend.position = "none")

aggregate(pokemon.dat$Total~pokemon.dat$Evolution, FUN=mean)
```

```{r}
pokemon.dat <- pokemon.dat %>% mutate(Legendary.Status = ifelse(Legendary == T, 1 ,0))
pokemon.dat$Legendary.Status <- as.factor(pokemon.dat$Legendary.Status)

leg.colors <- c("darkslategray2", "mediumturquoise")

ggplot(data = pokemon.dat, aes(x = Legendary.Status, y = Total)) +
  geom_violin(aes(fill = Legendary.Status)) +
  geom_jitter(width = 0.1, alpha = 0.2) +
  scale_fill_manual(values = leg.colors) +
  labs(title = "Pokemon Total Stats by Legendary Status",
       x = "Legendary Status",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5),
        legend.position = "none")
```


```{r}
tally(x = pokemon.dat)
# table(pokemon.dat$Type2)
```



$~$

#### Write out a statistical model to explore your research question.

$Total \ Stats_i \sim \ \displaystyle\sum_{j = 1}^{18}{\beta_j * Pokemon \ Type_i} + {\beta_{19} * Evolution_i} + \beta_{20} I_{Legendary, i} + \varepsilon_i$  

where $\varepsilon \sim N(0, \sigma^2)$, and $I_{Legendary}$ is an indicator variable for legendary status that is 1 when the pokemon is legendary and 0 otherwise.

$~$

#### Identify the parameters from the model & justify prior distributions.

$\beta_{bug}$ ~ N(400, 100), as we know the mean overall stats will be between 400 and 600 and believe bug types are kinda bad but we are not super sure about that. 

$\beta_{dark}$ ~ N(475, 100), as we believe dark types are kinda good but we are not super sure about that. 

$\beta_{dragon}$ ~ N(575, 50), as we believe dragon types are kinda the best and we are pretty sure about that. 

$\beta_{electric}$ ~ N(450, 75), as we believe electric types are pretty average but we are not so sure about that. 

$\beta_{fairy}$ ~ N(400, 100), as we believe fairy types kinda suck but we are not sure about that. 

$\beta_{fighting}$ ~ N(475, 75), as we believe fighting types are average but we are not so sure about that. 

$\beta_{fire}$ ~ N(475, 75), as we believe fire types are average but we are not so sure about that. 

$\beta_{flying}$ ~ N(500, 50), as we believe flying types are kinda the second best and we are pretty sure about that. 

$\beta_{ghost}$ ~ N(450, 75), as we believe ghost types are pretty average but we are not so sure about that. 

$\beta_{grass}$ ~ N(425, 75), as we believe grass types are below average but am not sure.

$\beta_{ground}$ ~ N(425, 75), as we believe grass types are below average but am not sure.

$\beta_{ice}$ ~ N(450, 75), as we believe ice types are pretty average but we are not so sure about that. 

$\beta_{normal}$ ~ N(400, 25), normal types are pretty much overall the worst, and we are sure about that. 

$\beta_{poison}$ ~ N(425, 75), as we believe poison types are below average but am not sure.

$\beta_{psychic}$ ~ N(475, 75), as we believe psychic types are average but am not sure.

$\beta_{rock}$ ~ N(425, 75), as we believe rock types are below average but am not sure.

$\beta_{steel}$ ~ N(450, 75), as we believe electric types are pretty average but we are not so sure about that. 

$\beta_{water}$ ~ N(450, 75), as we believe electric types are pretty average but we are not so sure about that. 


$\beta_{Evolution = 1}$

$\beta_{Evolution = 2}$

$\beta_{Evolution = 3}$

$\beta_{Evolution = 4}$



$\beta_{Legendary}$



$\sigma^2$ ~ Uniform(0, 1000^2), looking at the plot the variance is around 200 and it cannot be negative and we want the upper bound to be way greater than 200^2. 



$~$

#### Write JAGS code to fit this model

```{r}
geodude <- model.matrix(Total ~ Type + Evolution + Legendary.Status - 1, 
                        data = pokemon.dat)
```

```{r}
model_normal <- "model{
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(beta_bug * x1[i] + beta_dark * x2[i] + beta_dragon * x3[i] +
                 beta_electric * x4[i] + beta_fairy * x5[i] + beta_fighting * x6[i] +
                 beta_fire * x7[i] + beta_flying * x8[i] + beta_ghost * x9[i] +
                 beta_grass * x10[i] + beta_ground * x11[i] + beta_ice * x12[i] +
                 beta_normal * x13[i] + beta_poison * x14[i] + beta_psychic * x15[i] +
                 beta_rock * x16[i] + beta_steel * x17[i] + beta_water * x18[i] +
                 beta_gen1 * x19[i] + beta_gen2 * x20[i] + beta_gen3 * x21[i] +
                 beta_gen4 * x22[i] + beta_legend * x23[i],
                 1/(sigma^2)) 
  }
  
  # Priors
  beta_bug ~ dnorm(400, 1/100^2)
  beta_dark ~ dnorm(475, 1/100^2)
  beta_dragon ~ dnorm(575, 1/50^2)
  beta_electric ~ dnorm(450, 1/75^2)
  beta_fairy ~ dnorm(400, 1/100^2)
  beta_fighting ~ dnorm(475, 1/75^2)
  beta_fire ~ dnorm(475, 1/75^2)
  beta_flying ~ dnorm(500, 1/50^2)
  beta_ghost ~ dnorm(450, 1/75^2)
  beta_grass ~ dnorm(425, 1/75^2)
  beta_ground ~ dnorm(425, 1/75^2)
  beta_ice ~ dnorm(450, 1/75^2)
  beta_normal ~ dnorm(400, 1/25^2)
  beta_poison ~ dnorm(425, 1/75^2)
  beta_psychic ~ dnorm(475, 1/75^2)
  beta_rock ~ dnorm(425, 1/75^2)
  beta_steel ~ dnorm(450, 1/75^2)
  beta_water ~ dnorm(450, 1/75^2)
  
  beta_gen1 
  beta_gen2 
  beta_gen3
  beta_gen4 
  
  beta_legend
  
  sigma ~ dunif(0, 1000000)
}"

model <- jags.model(file = textConnection(model_normal), 
                   data = list(x1=geodude[,1], x2=geodude[,2], x3=geodude[,3],
                               x4=geodude[,4], x5=geodude[,5], x6=geodude[,6],
                               x7=geodude[,7], x8=geodude[,8], x9=geodude[,9],
                               x10=geodude[,10], x11=geodude[,11], x12=geodude[,12],
                               x13=geodude[,13], x14=geodude[,14], x15=geodude[,15],
                               x16=geodude[,16], x17=geodude[,17], x18=geodude[,18],
                               x19=geodude[,19], x20=geodude[,20], x21=geodude[,21],
                               x22=geodude[,22], x23=geodude[,23],
                               n = nrow(geodude)))

update(model, n.iter = 5000)

catchemall <- coda.samples(model, variable.names = c('beta_bug', 'beta_dark', 'beta_dragon', 'beta_electric', 'beta_fairy', 'beta_fighting', 'beta_fire', 'beta_flying', 'beta_ghost', 'beta_grass', 'beta_ground', 'beta_ice', 'beta_normal', 'beta_poison', 'beta_psychic', 'beta_rock', 'beta_steel', 'beta_water', 'sigma'), n.iter = 10000)
```


$~$

#### Summarize Parameters in the Model.

```{r}
summary(catchemall)
```



$~$

#### Graph Results

```{r}
want <- pokemon.dat[pokemon.dat$Type %in% c("Flying", "Bug", "Dragon", "Dark", "Rock"),]

want <- want %>%
  mutate(Type = factor(Type, 
                       levels = c("Flying", "Bug", "Dragon", "Dark", "Rock")))

want.CI <- data.frame(Type = c("Flying", "Bug", "Dragon", "Dark", "Rock"),
                      Mean.Est = c(412.18, 399.99, 398.78, 382.90, 382.85),
                      Upper.CI = c(478.31, 400.18, 427.48, 411.38, 406.96),
                      Lower.CI = c(346.81, 399.8, 370.78, 355.22, 359.37))

want.colors <- c("#00BF74", "#F8766D", "#D39200", "#E88526",  "#F564E3")

ggplot() +
  geom_violin(data = want, aes(x = Type, y = Total, fill = Type)) +
  geom_point(data = want.CI, aes(x = Type, y = Mean.Est)) +
  geom_errorbar(aes(x = Type, ymin = Lower.CI, ymax = Upper.CI), 
                data = want.CI, width = 0.3) +
  scale_fill_manual(values = want.colors) +
  labs(title = "Five Most Desirable Pokemon Types",
       x = "Pokemon Type",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.position = "none")
```


```{r}
dont <- pokemon.dat[pokemon.dat$Type %in% c("Fairy", "Ground", "Poison", "Normal", "Ice"),]

dont <- dont %>%
  mutate(Type = factor(Type, 
                       levels = c("Fairy", "Ground", "Poison", "Normal", "Ice")))

dont.CI <- data.frame(Type = c("Fairy", "Ground", "Poison", "Normal", "Ice"),
                      Mean.Est = c(346.37, 359.23, 354.53, 356.81, 358.30),
                      Upper.CI = c(383.45, 386.55, 383.37, 372.18, 389.14),
                      Lower.CI = c(309.49, 331.70, 325.85, 341.36, 326.57))

dont.colors <- c("#93AA00", "#00B9E3", "#AE87FF", "#619CFF", "#00ADFA")

ggplot() +
  geom_violin(data = dont, aes(x = Type, y = Total, fill = Type)) +
  geom_point(data = dont.CI, aes(x = Type, y = Mean.Est)) +
  geom_errorbar(aes(x = Type, ymin = Lower.CI, ymax = Upper.CI), 
                data = dont.CI, width = 0.3) +
  scale_fill_manual(values = dont.colors) +
  labs(title = "Five Least Desirable Pokemon Types",
       x = "Pokemon Type",
       y = "Estimated Total Stats") +
  ylim(0, 800) + 
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.position = "none")
```

$~$

#### Citations

```{r}
citation("ggplot2")
citation("rjags")
```

